---
title: "HW3"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rstan)
library(mnormt)
options(mc.cores = parallel::detectCores())
set.seed(02262021)
```

### Q1. GP Regression in 2D (12 points)

This question will continue the theme from the GP Demo on Friday February 12.

#### Simulate GP regression in 2D (4 points)

Simulate and visualize a GP in 2d with a covariate structure. To keep this simple, consider an additive relationship with the covariates, x1 and x2. While not required, take some time to explore the impacts of changing the total number of sampling locations as well as the parameters in the GP.

```{r}
phi <- .25
sigmasq <- 1
n <- 100
x1 <- runif(n)
x2 <- runif(n)
beta1 <- 10
beta2 <- -10
x <- cbind(x1,x2)
d <- sqrt(plgp::distance(x))
eps <- sqrt(.Machine$double.eps) 
H <- exp(-d/phi) + diag(eps, n) 
y <- rmnorm(1, x1 * beta1 + x2 * beta2,sigmasq * H)
reg_fig <- tibble(x1 = x1, x2 = x2, response = y) %>% ggplot(aes(y=x2, x=x1, color = response)) +
  theme_bw() + ggtitle('Random realization of a GP') +
  geom_point(size = 2) +  scale_colour_gradient2() + theme_dark()
reg_fig
```


#### Stan (4 points)

Now modify the Stan code to estimate this model in 2D. Report how well your code recovers the parameters (sigmasq, phi, beta0?, beta1?, beta2?) from your simulated dataset.

```{r, results = 'hide', message = F}
Reg_params <- stan("GP_reg_2d.stan", 
                  data=list(N = n, 
                            y = y,
                            x1 = x1,
                            dist = d,
                            x2 = x2))
```


```{r}
print(Reg_params, pars = c('beta1', 'beta2', 'phi', 'sigmasq'))
```

The model does a good job learning beta1 and beta 2. Phi and sigmasq are highly correlated and can give similar results for Sigma (despite not being the true value)


#### Make Predictions for a set of locations (4 points)

Using your results from the stan model, make predictions at a set of locations. The locations can be the entire surface, or just a handful of points in the space. Include these results in a figure.


```{r}
beta1_hat <- summary(Reg_params, pars = 'beta1')$summary[1]
beta2_hat <- summary(Reg_params, pars = 'beta2')$summary[1]
sigmasq_hat <- summary(Reg_params, pars = 'sigmasq')$summary[1]
phi_hat <- summary(Reg_params, pars = 'phi')$summary[1]
npred <- 200
x1new <- runif(npred)
x2new <- runif(npred)
xnew <-cbind(x1new, x2new)
d1 <- sqrt(plgp::distance(xnew))
Sigma11 <- sigmasq_hat * exp(-d1/phi_hat) + diag(eps, npred) 
d12 <- sqrt(plgp::distance(xnew,x))
Sigma12 <- sigmasq_hat * exp(-d12/phi_hat) 
Sigma22 <- sigmasq_hat * exp(-d/phi_hat) + diag(eps, n) 

mu_1given2 <- x1new * beta1_hat + x2new * beta2_hat + Sigma12 %*% solve(Sigma22) %*% matrix(y - x1 * beta1_hat - x2 * beta2_hat, nrow = length(y), ncol = 1)
Sigma_1given2 <- Sigma11 - Sigma12 %*% solve(Sigma22) %*% t(Sigma12)
```

\vfill

```{r, echo = F}
mean_vals <- tibble(y_mean = mu_1given2, x1new = x1new, x2new = x2new)
data_and_mean <- reg_fig + 
    geom_point(aes(y = x2new, x = x1new), inherit.aes = F, data = mean_vals, size = 4) + 
  geom_point(aes(y = x2new, x = x1new, color = y_mean), inherit.aes = F, data = mean_vals, size = 2) + 
  geom_point(size = 2) + ggtitle("Observed Data + Conditional Mean")
data_and_mean + labs(caption = 'Predicted points are outlined in black')
```
